---
title: "Group 5: NYC Citibike Data Analysis"
author: "Anusha Muddapati, Harshith Sesham, Deekshit Vedula, Tejeshwine Viswanathan, Loren Young"
date: "4/21/2022"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#rmarkdown::render

library(knitr) 
library(RMySQL)
library(stringr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(geosphere)
library(sf)
library(broom)
library(sp)
library(grid)
library(stringr)
library(lubridate)

nyc_bike_data <- read.csv("202201-citibike-NYC.csv")
nyc_bike_data <- nyc_bike_data %>% drop_na()

temp_data <- read.csv("DailyWeatherDataNYCJan2022.csv")

# insert code for plots here
started_at <- nyc_bike_data$started_at
          
ended_at <- nyc_bike_data$ended_at
          
started_at_df <- as.data.frame(started_at)
ended_at_df <- as.data.frame(ended_at)
          
started_at_date <- as.Date(nyc_bike_data$started_at, format="%Y-%m-%d")
ended_at_date <- as.Date(nyc_bike_data$ended_at, format="%Y-%m-%d")
          
nyc_bike_data$started_at_date <- started_at_date
nyc_bike_data$ended_at_date <- ended_at_date

no_of_bikes <- count(nyc_bike_data %>% group_by(started_at_date))
          
dates <- format(as.Date(temp_data$datetime, format="%Y-%m-%d"))
          
temp <- temp_data$temp
min_temp <- temp_data$tempmin
max_temp <- temp_data$tempmax

member_casual_counts <- count(nyc_bike_data %>% 
                              group_by(started_at_date, member_casual))
member_casual_count <- count(nyc_bike_data %>% group_by(started_at_date, member_casual))

stations_count <- count(nyc_bike_data %>% group_by(start_station_name))
top_5_stations <- stations_count[order(-stations_count$n),][1:5,]$start_station_name
          
top_5_stations_performance <- nyc_bike_data %>% 
                                group_by(started_at_date, start_station_name) %>%
                                filter(start_station_name %in% top_5_stations)


total_bike_counts_and_weather <- count(nyc_bike_data %>% 
                                       group_by(started_at_date))
precipitation_data <- temp_data$precip
total_bike_counts_and_weather$precip <- precipitation_data

total_bike_counts_and_weather$temp <- temp
          
          
icon_data <- temp_data$icon
          
total_bike_counts_and_weather$icon <- icon_data

distance_travelled_data <- nyc_bike_data %>%
  mutate(distance_travelled = distHaversine(cbind(start_lat, start_lng), cbind(end_lat, end_lng))) %>%
  select(started_at_date, ended_at_date, member_casual, distance_travelled) %>% drop_na()

distance_travelled_data$distance_travelled <- distance_travelled_data$distance_travelled/1000

mean_distance_travelled_by_date <- aggregate(distance_travelled_data$distance_travelled, 
                                                  list(distance_travelled_data$started_at_date), FUN=mean)
          
          
          
max_distance_travlled_by_date <- aggregate(distance_travelled_data$distance_travelled, 
                                                     list(distance_travelled_data$started_at_date), FUN=max)
          
           
avg_co2_emission_per_km = 0.255 # in kgs
          
total_distance_travelled_by_date <- aggregate(distance_travelled_data$distance_travelled, 
                                                    list(distance_travelled_data$started_at_date), FUN=sum)

co2_reduction_per_day <- total_distance_travelled_by_date$x * 0.255
          
distance_data_by_date <- as.data.frame(mean_distance_travelled_by_date$Group.1)
colnames(distance_data_by_date)[1] <- "Dates"
          
distance_data_by_date$mean_dist <- mean_distance_travelled_by_date$x
distance_data_by_date$max_dist <- max_distance_travlled_by_date$x
distance_data_by_date$total_dist <- total_distance_travelled_by_date$x
distance_data_by_date$total_co2_reduction <- co2_reduction_per_day

#########
bike_data <- read.csv("202201-citibike-NYC.csv")

weather_data <- read.csv("DailyWeatherDataNYCJan2022.csv")
#bike_data$started_at
starting_time <- format(as.POSIXct(bike_data$started_at), format = "%H:%M:%S")

tod <- cut(chron::times(starting_time) , breaks = (1/24) * c(0,5,11,16,19,24))
bike_data$part_of_day <- c("Night","Morning","Day","Evening","Night")[as.numeric(tod)]
  
#Duration of ride and converting the date column to datetime
bike_data$duration_of_ride <- difftime(bike_data$ended_at,bike_data$started_at, 
                                       units = 'hours')
bike_data <- bike_data %>% rename(datetime = started_at)
bike_data$datetime <- as.Date(bike_data$datetime)

# total num hours per day:
total_hours_per_day <- bike_data %>% group_by(as.Date(datetime)) %>% 
  summarize(num_hours = sum(duration_of_ride))

weather_data$duration <- total_hours_per_day$num_hours
weather_data %>% select(datetime, duration)
#weekdays of dates
weather_data$weekdays <- wday(weather_data$datetime, label=TRUE, abbr=FALSE)
names(bike_data)
names(weather_data)

#num of rides per day and rename columns to have apt names
riders_per_day <- bike_data %>% group_by(as.Date(bike_data$datetime)) %>% count()
names(riders_per_day)

riders_per_day <- riders_per_day %>% rename(date_of_ride = `as.Date(bike_data$datetime)`,
                          num_riders = n )


weather_data$datetime <- as.Date(weather_data$datetime)
joint_dataset <- inner_join(bike_data, weather_data, by = 'datetime')

joint_dataset[is.na(joint_dataset$part_of_day),]$part_of_day <- 'Night'

temperatureColor <- "#cc9900"
priceColor <- "#CC3333"


library(ggpubr)
#install.packages("ggpmisc")
library(ggpmisc)

library(knitr)
library(tidyverse)
library(forcats)
library(gridExtra)
library(jsonlite)
library(gtable) 
library(grid)
library(latex2exp)
library(gridBase)
library(nnet)
library(magrittr)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
"#F0E442", "#0072B2", "#D55E00", "#CC79A7")
theme1 <- theme_bw() +
theme(axis.text = element_text(size = 8, colour = "#003300"),
axis.title = element_text(size = 15, colour = "brown"),
legend.title = element_text(size = 15, colour = "#660000"),
legend.text = element_text(size = 13, colour = "darkred"),
title = element_text(size = 15, colour = "#2f2f63"),
axis.ticks = element_line(colour = "#6b3447"),
plot.caption = element_text(size = 8, colour = "#2f2f63"),
plot.subtitle = element_text(size = 10, colour = "#2f2f63"))


joint_dataset[joint_dataset$conditions == "Snow, Rain, Freezing Drizzle/Freezing Rain, Partially cloudy",]$conditions <- "Freezing Drizzle/Freezing Rain"
legend_title = "Membership status"


  
  weather_data$average_duration <- weather_data$duration/riders_per_day$num_riders

  
  sorted_riders_per_day <- riders_per_day[order(riders_per_day$date_of_ride),]

weather_data_sorted <- weather_data[order(as.Date(weather_data$datetime)),]

weather_data_sorted$datetime
sorted_riders_per_day$date_of_ride


library(magrittr)
library(geodist)
library(hms)
library(grid)
library(corrplot)
library(tidyr)

nyc_bike <- read.csv("202201-citibike-NYC.csv") %>% data.frame()
nyc_weather <-read.csv("DailyWeatherDataNYCJan2022.csv") %>% 
  data.frame()

 nyc_bike$started_at %>% str_sub(12,19) %>% as_hms()-> nyc_bike$time_start
nyc_bike$ended_at %>% str_sub(12,19) %>% as_hms() -> nyc_bike$time_end
nyc_bike$duration <- round(nyc_bike$time_end-nyc_bike$time_start,3)

#converting duration from seconds to hours to store in duration_hours
nyc_bike$duration %>% as.numeric() ->nyc_bike$duration_hours
nyc_bike$duration_hours <- round(nyc_bike$duration_hours/3600,5)
###################################################################
#fixing weekdays to each of the days by extracting date
nyc_bike$started_at %>% str_sub(1,10) -> nyc_bike$datetime
nyc_bike$weekday <- weekdays(as.Date(nyc_bike$datetime))
#drop na values
nyc_bike %>% drop_na() -> nyc_bike
#View(nyc_bike)

#Merge the nyc_bike and nyc_weather data as nyc
nyc_bike %>% merge(nyc_weather,by="datetime") %>%  
  as.data.frame()-> nyc
#calculate count of bikers per day as "n"
nyc %>% count(datetime) -> cnt
nyc %>% right_join(cnt,by=c("datetime")) %>% 
  as.data.frame() ->nyc
#View(nyc)

#Calculate the distance traveled and speed in mph.

nyc$dist_miles <- geodist_vec(
  x1 = nyc$start_lng
  , y1 = nyc$start_lat
  , x2 = nyc$end_lng
  , y2 = nyc$end_lat
  , paired = TRUE
  , measure = "haversine"
)
#convert distance from meters to miles
nyc %>% transmute(dist_miles=dist_miles/1609) %>% round(5)-> nyc$dist_miles

#calculate speed mph
nyc$speed_mph <- round(nyc$dist_miles/nyc$duration_hours,5)
nyc %>% filter(duration_hours>0) -> nyc

# names(nyc)


  rm(mean)
aggregate(unlist(speed_mph)~datetime, nyc, FUN=mean) %>% 
  as.data.frame() -> avg_speed
#temp
aggregate(temp~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_temp
#precip
aggregate(precip~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_precip
#humidity
aggregate(humidity~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_humidity
#windspeed
aggregate(windspeed~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_windspeed
#sealevelpressure
aggregate(sealevelpressure~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_sealevelpressure
#cloudcover
aggregate(cloudcover~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_cloudcover
#solarradiation
aggregate(solarradiation~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_solarradiation
#visibility
aggregate(visibility~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_visibility
names(nyc)
#rename columns
avg_speed$speed <- avg_speed$`unlist(speed_mph)`
avg_speed = subset(avg_speed, select = -c(`unlist(speed_mph)`) )

#Join avg speed with all the avg weather conditions
avg_speed %>% inner_join(avg_temp,by="datetime") %>% 
  as.data.frame() %>% inner_join(avg_precip,by="datetime") %>% 
  as.data.frame() %>% inner_join(avg_humidity,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_windspeed,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_sealevelpressure,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_cloudcover,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_solarradiation,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_visibility,by="datetime")-> avg_speed

avg_speed$weekday <- weekdays(as.Date(avg_speed$datetime))
#View(avg_speed)

#DO THE FREAKIN PLOTS!!!!!!!!!!

drop <- c("datetime","weekday")
avg_speed_subset <-  avg_speed[,!(names(avg_speed) %in% drop)]
#View(avg_speed_subset)

```

## I. Introduction

- This project analyzes the correlation and influences that weather has on public bicycling in the New York City (NYC) metropolitan area. 
- Using weather data taken from the National Oceanic and Atmospheric Administration (NOAA) for the month of January, 2022, these influences were examined on bicycle riders under the Citi BikeShare community.
- The following analysis allows for the understanding and interpretation of how weather impacts commuting via bicycle for a large American city in contemporary times.


## II. Data-set

- Citi BikeShare is a rental company which allows users to rent bicycles. A rider unlocks a bike and pays per time or distance that they use the bike. After a ride is ended at a designated BikeShare location, thus ending the ride, the data is recorded in the BikeShare database. 
- Parameters recorded include the type of bike used, the start and end times, and the start and end geographical coordinates and the start and end stations. 
- The weather data includes the type of cloud cover, precipitation, and temperature for different times of day during the month of January, 2022. 
- Both data-sets thus align in their respective time-frames, with the BikeShare data-set recording over one million events for the month. 

	
## III. Analysis

- The premise of the analysis is to understand the correlation between several weather variables and BikeShare riders for the NYC area. 
- Both data-sets were accumulated into a SQL database for easier accessibility of each parameter. 
- Several weather related parameters were plotted against the types of bikes and number of riders for each day of the month. 
- Regression analysis was employed to determine the correlations between parameters in the two data-sets to observe trends in the data. 


## Number of bikers

```{r plot1}

renderPlot({
  
ggplot(member_casual_counts, aes(x=started_at_date, y=n, group=member_casual)) +
       geom_line(aes(color=member_casual)) +
       geom_point(aes(color=member_casual)) +
       ggtitle("Line graph of number of bikes") + xlab("Date") +ylab("Number of bikes")
  
})
```



```{r plot2}
# ## Number of bikers by member type
# 
# renderPlot({
# ggplot(data=member_casual_count, aes(x=started_at_date, y=n, fill=member_casual))+
#        geom_bar(stat='identity') + xlab("Date") + ylab("Number of bikes") +
#        labs(title="Number of bikes by member type")
# 
# })


```




```{r}

# # Number of bikes by membership (box)
# renderPlot({ggplot(member_casual_counts, aes(x=member_casual, y=n, fill=member_casual)) +
#        geom_boxplot(alpha=0.8) +
#        theme(legend.position = "none") + ggtitle("Box plot of bike membership") +
#        xlab("") + ylab("Number of bikes")
# 
# })
```



## Top 5 best performing stations

```{r}

renderPlot({
          
plot <- ggplot(top_5_stations_performance) +
               aes(x=started_at_date, fill=start_station_name) + geom_bar() +
               ggtitle("Top 5 stations' performance") +
               ylab("Number of bikes") + xlab("Date")
          
plot <- plot +
          facet_wrap(~start_station_name, ncol=3, scales="free") +
          theme(legend.position = "none")
          
plot
})
```


## Average CO2 emission reduced

```{r}

# Avg CO2 emission reduced
renderPlot({ggplot(data=distance_data_by_date, aes(x=Dates, y=total_co2_reduction)) +
            geom_bar(stat="identity", fill="steelblue") + 
            xlab("Date") + ylab("Average CO2 emisison reduced in Kgs") +
            ggtitle("Average CO2 emission reduced")
})
```

## Density plot of distance travelled (by membership)

```{r}

# Density plot of distance travelled by membership
renderPlot({ggplot(distance_travelled_data, aes(x=distance_travelled, fill=member_casual)) +
       geom_density(alpha=0.3) + xlab("Distance travelled") + ylab("Density") + 
       ggtitle("Density plot of distances travelled")
})
```


## Number of bikes with Temperature (Min and Max)

```{r}
# Number of bikes with Temperature
renderPlot({plot(no_of_bikes$started_at_date, no_of_bikes$n, type="l", lty=1,
     lwd=2, col=3, main="Number of Bikes with Temperature", 
     xlab="Date", ylab="Number of bikes")
axis(side=2, col=3)
          
par(new=TRUE)
          
plot(no_of_bikes$started_at_date, temp, type="l", col = 6, 
     axes = FALSE, xlab = "", ylab = "")
arrows(x0=no_of_bikes$started_at_date, y0=min_temp, 
       x1=no_of_bikes$started_at_date, y1=max_temp, code=3, angle=90, 
       length=0.05, col="blue", lwd=0.2)
axis(side = 4, at = pretty(range(temp)), col=6)
mtext("Temperature", side=4)
})
```





```{r}
# ## Number of bikes with precipitation
# # Number of bikes vs precipitation
# renderPlot({
#           
# plot(total_bike_counts_and_weather$started_at_date, total_bike_counts_and_weather$n, type="l", lty=1,
#      lwd=2, col=3, main="Number of Bikes with Precipitation", 
#      xlab="Date", ylab="Number of bikes")
# axis(side=2, col=3)
#                                                   
# par(new=TRUE)
#           
# plot(total_bike_counts_and_weather$started_at_date, total_bike_counts_and_weather$precip, 
#      type="l", col = 5, 
#      axes = FALSE, xlab = "", ylab = "")
# axis(side = 4, at = pretty(range(total_bike_counts_and_weather$precip)), col=5)
# mtext("Precipitation", side=4)
# })
```

## Number of bikers by weather type

```{r}

# Box plot for number of rides by type of weather
# total_bike_counts_and_weather <- count(nyc_bike_data %>% 
#                                        group_by(started_at_date))
#           
renderPlot({
          
ggplot(total_bike_counts_and_weather, aes(x=icon, y=n, fill=icon)) +
       geom_boxplot(alpha=0.8) +
       theme(legend.position = "none") + ggtitle("Number of bikes based on weather type") +
       xlab("") + ylab("Number of bikes")
})
```




```{r}
# ## Distance travelled
# # Box plot of distances travelled
# renderPlot({
#           
# ggplot(distance_travelled_data, 
#        aes(x=member_casual, y=distance_travelled, fill=member_casual)) + 
#        geom_boxplot() + xlab("") + ylab("Distance travelled in kms")
# })
```



```{r}
## Density plot of distance travelled
# # Density plot of distance travelled
# renderPlot({ggplot(distance_travelled_data, aes(x=distance_travelled)) +
#        geom_density(color="darkgreen", fill="lightgreen", alpha=0.8) + 
#        xlab("Distance travelled") + ylab("Density") + ggtitle("Density plot of distances travelled")
# })
```




```{r}
# 
# # Avg and Total Distance travelled ## Average and Maximum distance travelled
# renderPlot({
#           
# p1 <- ggplot(data=distance_data_by_date, aes(x=Dates, y=mean_dist)) +
#             geom_line(color="darkred") + xlab("Date") + ylab("Average Distance travelled in kms") +
#             ggtitle("Average Distance travelled by a rider")
#           
# p2 <- ggplot(data=distance_data_by_date, aes(x=Dates, y=max_dist)) +
#             geom_line(color="steelblue") + xlab("Date") + ylab("Maximum Distance travelled in kms") +
#             ggtitle("Maximum Distance travelled by a rider")
#           
# ggarrange(p1, p2, nrow = 1)
# })

```





```{r}
# ##  Distribution of rider type Vs Temperature
##ggplot(joint_dataset, aes(x=datetime, fill = member_casual, color=member_casual)) +
#   geom_histogram(alpha=0.5, bins = 30)+
#   geom_line(aes(y=temp*1000), color = 'darkmagenta', size = 1) +
#   geom_point(aes(y=temp*1000), color = 'darkmagenta')+
#   scale_y_continuous(
#     name = "Number of rides",
#     sec.axis = sec_axis(~./1000, name="Temperature")
#   )+
#   labs(x = "Date") +
#   scale_x_date(date_labels="%d %b",date_breaks  ="1 day")+
#   theme(axis.text.x = element_text(angle=90)) +
#   ggtitle("Distribution of rider type Vs Temperature")
#   #coord_cartesian(xlim=c(as.Date("31 Dec"),as.Date("31 Jan")))
  

```


## Average duration of rides daily against Temperature

```{r fig.alig="center"}

renderPlot({
  
  
ggplot(weather_data) + aes(x = datetime) +
    geom_point(aes(y = average_duration*60), size=2, color="red") + 
  geom_line(aes(y = average_duration*60), color="red") + 
  geom_line(aes(y = temp), color = "blue") +
  geom_point(aes(y = temp), size=2, color="blue") + scale_y_continuous(
    name = "Average duration(mins) ridden daily",
    sec.axis = sec_axis(~., name="Temperature")
  ) +
  theme1 + 
  facet_wrap(~ weekdays, ncol=3, scales="free" )  + 
  theme(strip.text.x = element_text(size = 8)) +
  theme(axis.text.x = element_text(angle=90, hjust=1,size=8)) +
  theme(axis.text.y = element_text(size=8)) + 
  theme(axis.title=element_text(size=12)) +
  theme(plot.title=element_text(size=10)) +
  scale_fill_manual(values =cbPalette, legend_title) +
  theme(
    axis.title.y = element_text(color = "red", size=12),
    axis.title.y.right = element_text(color = "blue", size=12),
    axis.title.x =  element_text()
   # axis.text.x = element_text(angle = 90)
  ) +
  labs(x = "Date") +
  scale_x_date(date_labels="%d %b",date_breaks  ="1 day") 
 # ggtitle("Average duration of rides daily against Temperature across Weekdays")
}, height = 400, width = 800)

```


## Number of Riders by Membership status and Time of day

```{r}


renderPlot({

  ggplot(joint_dataset) +
    geom_bar(position = "dodge", aes(x = part_of_day, fill = member_casual)) + 
  theme1 + 
  facet_wrap(~ weekdays,ncol=3, scales="free" )  + 
  theme(strip.text.x = element_text(size = 8)) +
  theme(axis.text.x = element_text(angle=45, hjust=1,size=8)) +
  theme(axis.text.y = element_text(size=8)) + 
  theme(axis.title=element_text(size=12)) +
  theme(plot.title=element_text(size=10)) +
  labs(x="Time of day",y="Number of riders") + 
  scale_fill_manual(values =cbPalette, legend_title)
}, height = 400, width = 800)

#title="Number of Riders by Membership status and Part of day
      # Based on Weekdays",
```




```{r}
## Linear Regression of Number of rides per day Vs Temperature on that day
#LINEAR REGRESSION PLOT
# renderPlot({
#   
# ggplot() +aes(weather_data$temp, riders_per_day$num_riders) +
#   geom_point(color = 'red', size = 3) +
#   geom_smooth(method='lm', color = 'darkgreen') +
#   stat_regline_equation(label.x.npc = "middle")+
#   stat_fit_glance(method = 'lm',
#                        geom = 'text',
#                        aes(label = paste("P-value = ", signif(..p.value.., digits = 4), sep = "")),
#                   label.x.npc = 'right', label.y.npc = 0.65, size = 5)+
#   labs(y='Number of rides per day', x='Temperature') +
#   theme(plot.title = element_text(hjust=0.5, size=12, face='bold'),
#         axis.title = element_text(size=12))
#  # theme_bw()
# })

```


## Linear Regression: Number of Rides ~ Weather conditions

```{r}

library(shiny)

  nyc_bike <- read.csv("202201-citibike-NYC.csv") %>% data.frame()
nyc_weather <-read.csv("DailyWeatherDataNYCJan2022.csv") %>% 
  data.frame()

 nyc_bike$started_at %>% str_sub(12,19) %>% as_hms()-> nyc_bike$time_start
nyc_bike$ended_at %>% str_sub(12,19) %>% as_hms() -> nyc_bike$time_end
nyc_bike$duration <- round(nyc_bike$time_end-nyc_bike$time_start,3)

#converting duration from seconds to hours to store in duration_hours
nyc_bike$duration %>% as.numeric() ->nyc_bike$duration_hours
nyc_bike$duration_hours <- round(nyc_bike$duration_hours/3600,5)
###################################################################
#fixing weekdays to each of the days by extracting date
nyc_bike$started_at %>% str_sub(1,10) -> nyc_bike$datetime
nyc_bike$weekday <- weekdays(as.Date(nyc_bike$datetime))
#drop na values
nyc_bike %>% drop_na() -> nyc_bike
#View(nyc_bike)

#Merge the nyc_bike and nyc_weather data as nyc
nyc_bike %>% merge(nyc_weather,by="datetime") %>%  
  as.data.frame()-> nyc
#calculate count of bikers per day as "n"
nyc %>% count(datetime) -> cnt
nyc %>% right_join(cnt,by=c("datetime")) %>% 
  as.data.frame() ->nyc
#View(nyc)

#Calculate the distance traveled and speed in mph.

nyc$dist_miles <- geodist_vec(
  x1 = nyc$start_lng
  , y1 = nyc$start_lat
  , x2 = nyc$end_lng
  , y2 = nyc$end_lat
  , paired = TRUE
  , measure = "haversine"
)
#convert distance from meters to miles
nyc %>% transmute(dist_miles=dist_miles/1609) %>% round(5)-> nyc$dist_miles

#calculate speed mph
nyc$speed_mph <- round(nyc$dist_miles/nyc$duration_hours,5)
nyc %>% filter(duration_hours>0) -> nyc

# names(nyc)


#  rm(mean)
aggregate(unlist(speed_mph)~datetime, nyc, FUN=mean) %>% 
  as.data.frame() -> avg_speed
#temp
aggregate(temp~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_temp
#precip
aggregate(precip~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_precip
#humidity
aggregate(humidity~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_humidity
#windspeed
aggregate(windspeed~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_windspeed
#sealevelpressure
aggregate(sealevelpressure~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_sealevelpressure
#cloudcover
aggregate(cloudcover~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_cloudcover
#solarradiation
aggregate(solarradiation~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_solarradiation
#visibility
aggregate(visibility~datetime,nyc,FUN=mean) %>% 
  as.data.frame() -> avg_visibility
#names(nyc)
#rename columns
avg_speed$speed <- avg_speed$`unlist(speed_mph)`
avg_speed = subset(avg_speed, select = -c(`unlist(speed_mph)`) )

#Join avg speed with all the avg weather conditions
avg_speed %>% inner_join(avg_temp,by="datetime") %>% 
  as.data.frame() %>% inner_join(avg_precip,by="datetime") %>% 
  as.data.frame() %>% inner_join(avg_humidity,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_windspeed,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_sealevelpressure,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_cloudcover,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_solarradiation,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_visibility,by="datetime")-> avg_speed

avg_speed$weekday <- weekdays(as.Date(avg_speed$datetime))


drop <- c("datetime","weekday")
avg_speed_subset <-  avg_speed[,!(names(avg_speed) %in% drop)]

drop <- c("datetime")
avg_speed_subset2 <-  avg_speed[,!(names(avg_speed) %in% drop)]
required_vars <- avg_speed_subset2 %>%
  gather(-speed, -weekday, key = "var", value = "value")
required_lst <- names(table(required_vars$var))


aggregate(n~datetime, nyc, FUN=mean) %>% 
  as.data.frame() -> cnt
cnt %>% inner_join(avg_temp,by="datetime") %>% 
  as.data.frame() %>% inner_join(avg_precip,by="datetime") %>% 
  as.data.frame() %>% inner_join(avg_humidity,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_windspeed,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_sealevelpressure,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_cloudcover,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_solarradiation,by="datetime")%>% 
  as.data.frame() %>% inner_join(avg_visibility,by="datetime")-> cnt

cnt$weekday <- weekdays(as.Date(cnt$datetime))


drop <- c("datetime")
#names(cnt)
cnt <-  cnt[,!(names(avg_speed) %in% drop)]
cnt_bikers_weather <- cnt %>%
  gather(-n, -weekday, key = "var", value = "value") 



draw_plot2 <- function(var_to_filter_by) {
  
  cnt_bikers_weather %>%
    filter(var == !!var_to_filter_by) %>%
    ggplot(aes(x = value, y = n)) +
  geom_point(size = 3, color = 'red') +
  geom_smooth(method='lm', color = 'darkgreen') +
  stat_regline_equation(label.x.npc = "middle")+
  stat_fit_glance(method = 'lm',
                       geom = 'text',
                       aes(label = paste("P-value = ", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = 'right', label.y.npc = 0.65, size = 3)+
  labs(y='Number of rides per day', x=as.character(var_to_filter_by), title=paste0('Linear Regression of Number of rides per day Vs ', as.character(var_to_filter_by))) +
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'))
  
}

ui2 <- fluidPage(
  inputPanel(
    selectInput(
      "var",
      label = "Select Weather Condition",
      choices = required_lst
    )
  ),
  plotOutput('LRplot')
)

server2 <- function(input, output) {
  output$LRplot <- renderPlot(draw_plot2(input$var), height = 325, width = 750)
}

shinyApp(ui = ui2, server = server2)



```

## Correlation plot of Average Speed of Bikers per day against Weather Conditions

```{r}
#CORRELATION PLOT
renderPlot({
  
  M=cor(avg_speed_subset)
corrplot(M, method = 'square', order = 'FPC', type = 'lower', 
         tl.col = "black", diag = FALSE)
}, height = 360, width = 700)

```

## Speed By Temperature

```{r}
#TEMP VS SPEED ACC TO PRECIP
renderPlot({tempvsspeed <- avg_speed %>% 
  ggplot(aes(x = speed, 
             y = temp,color=precip)) +
  geom_point(show.legend = TRUE, size = 3) +
  facet_grid(~weekday) +
  theme(axis.title = element_text()) +
  ylab("Temperature (°F)") +
  xlab("Speed (mph)") +
  ggtitle("Speed By Temperature")+
  theme(plot.title=element_text(hjust=0.5))+
  theme(axis.text.x = element_text(face="bold", size=8),
        axis.text.y = element_text(face="bold", size=8))
tempvsspeed

})

```




```{r}
#SPEED VS ALL CONDITIONS
## Average Speed of Bikers per day against Weather Conditions
# renderPlot({drop <- c("datetime")
# avg_speed_subset2 <-  avg_speed[,!(names(avg_speed) %in% drop)]
# avg_speed_subset2 %>%
#   gather(-speed, -weekday, key = "var", value = "value") %>% 
#   ggplot(aes(x = value, y = speed, color = weekday)) +
#   geom_point() +
#   facet_wrap(~ var, scales = "free") +
#   theme_bw()+
#   theme(axis.title = element_text()) +
#   xlab("Different weather conditions") +
#   ylab("Average Speed of bikers (mph)") +
#   ggtitle("Impact of Weather conditions on average speed")+
#   theme(plot.title = element_text(face = "bold",hjust = 0.5,color="grey20"))
# 
# })

```




```{r}

## Biker Density vs Weather Conditions
# renderPlot({
#   
# aggregate(n~datetime, nyc, FUN=mean) %>% 
#   as.data.frame() -> cnt
# cnt %>% inner_join(avg_temp,by="datetime") %>% 
#   as.data.frame() %>% inner_join(avg_precip,by="datetime") %>% 
#   as.data.frame() %>% inner_join(avg_humidity,by="datetime")%>% 
#   as.data.frame() %>% inner_join(avg_windspeed,by="datetime")%>% 
#   as.data.frame() %>% inner_join(avg_sealevelpressure,by="datetime")%>% 
#   as.data.frame() %>% inner_join(avg_cloudcover,by="datetime")%>% 
#   as.data.frame() %>% inner_join(avg_solarradiation,by="datetime")%>% 
#   as.data.frame() %>% inner_join(avg_visibility,by="datetime")-> cnt
# 
# cnt$weekday <- weekdays(as.Date(cnt$datetime))
# 
#   drop <- c("datetime")
# names(cnt)
# cnt <-  cnt[,!(names(avg_speed) %in% drop)]
# cnt %>%
#   gather(-n, -weekday, key = "var", value = "value") %>% 
#   ggplot(aes(x = value, y = n, color = weekday)) +
#   geom_point() +
#   facet_wrap(~ var, scales = "free") +
#   theme_bw()+
#   theme(axis.title = element_text()) +
#   xlab("Different weather conditions") +
#   ylab("Number of bikers") +
#   ggtitle("Impact of Weather conditions on biker density")+
#   theme(plot.title = element_text(face = "bold",hjust = 0.5,color="grey20"))
# })

```



```{r}
#SIMPLE PLOTS SUCH AS WEEKDAY COUNT BASED ON MEMBERSHIP AND BIKE
## Biker density over the week (on a Membership basis)
# renderPlot({ggplot(nyc,aes(x=reorder(weekday,-n,sum),y=n,fill=member_casual))+
#   theme_bw()+geom_col()+
#   labs(x='Day',y='Number of bikers',title='Biker density over the week based on membership')
# })


```




```{r}

# renderPlot({ggRentalsByTemp <- nyc %>%
#   ggplot(aes(y = n,
#                  x = temp,
#                  color = weekday)) +
#   geom_point(show.legend = TRUE) +
#   theme(legend.position=c(0.95,0.95))+
#   theme(legend.title = element_text(face = "bold"))+
#   theme(legend.text = element_text(size = 8, colour = "black"))+
#   theme(legend.key = element_rect(fill = "white", colour = "grey64"))+
#   facet_grid(~weekday) +
#   scale_color_brewer(palette = "Dark2") +
#   theme_bw() +
#   theme(axis.title = element_text()) +
#   ylab("Bike Rentals") +
#   xlab("Temperature (°F)") +
#   ggtitle("Bike Rental Volume By Temperature")+
#   theme(plot.title=element_text(hjust=0.5))+
#   guides(color='none')
# ############################################################################
# ggRentalsByhumidity <- nyc_corrected %>%
#   ggplot(aes(y = n,
#                  x = humidity,
#                  color = weekday)) +
#   geom_point(show.legend = TRUE) +
#   theme(legend.position=c(0.95,0.95))+
#   theme(legend.title = element_text(face = "bold"))+
#   theme(legend.text = element_text(size = 8, colour = "black"))+
#   theme(legend.key = element_rect(fill = "white", colour = "grey64"))+
#   facet_grid(~weekday) +
#   scale_color_brewer(palette = "Dark2") +
#   theme_bw() +
#   theme(axis.title = element_text()) +
#   ylab("Bike Rentals") +
#   xlab("Humidity") +
#   ggtitle("Bike Rental Volume By Humidity")+
#   theme(plot.title=element_text(hjust=0.5))+
#   guides(color='none')
# #############################################################################
# ggRentalsBycloudcover <- nyc_corrected%>%
#   ggplot(aes(y = n,
#                  x = cloudcover,
#                  color = weekday)) +
#   geom_point(show.legend = TRUE) +
#   theme(legend.position=c(0.95,0.95))+
#   theme(legend.title = element_text(face = "bold"))+
#   theme(legend.text = element_text(size = 8, colour = "black"))+
#   theme(legend.key = element_rect(fill = "white", colour = "grey64"))+
#   facet_grid(~weekday) +
#   scale_color_brewer(palette = "Dark2") +
#   theme_bw() +
#   theme(axis.title = element_text()) +
#   ylab("Bike Rentals") +
#   xlab("Cloud cover") +
#   ggtitle("Bike Rental Volume By Cloud cover")+
#   theme(plot.title=element_text(hjust=0.5))+
#   guides(color='none')
# #############################################################################
# ggRentalsByprecip <- nyc %>%
#   ggplot(aes(y = n,
#                  x = precip,
#                  color = weekday
#              )) +
#   geom_point(show.legend = TRUE) +
#   theme(legend.position=c(0.95,0.95))+
#   theme(legend.title = element_text(face = "bold"))+
#   theme(legend.text = element_text(size = 8, colour = "black"))+
#   theme(legend.key = element_rect(fill = "white", colour = "grey64"))+
#   facet_grid(~weekday) +
#   scale_color_brewer(palette = "Dark2") +
#   theme_bw() +
#   theme(axis.title = element_text()) +
#   ylab("Bike Rentals") +
#   xlab("Rain") +
#   ggtitle("Bike Rental Volume By Rain")+
#   theme(plot.title=element_text(hjust=0.5))+
#   guides(color='none')
# ############################################################################
# ggRentalsByvisibility <- nyc %>%
#   ggplot(aes(y = n,
#                  x = visibility,
#                  color = weekday))+
#   geom_point(show.legend = TRUE) +
#   theme(legend.position=c(0.95,0.95))+
#   theme(legend.title = element_text(face = "bold"))+
#   theme(legend.text = element_text(size = 8, colour = "black"))+
#   theme(legend.key = element_rect(fill = "white", colour = "grey64"))+
#   facet_grid(~weekday) +
#   scale_color_brewer(palette = "Dark2") +
#   theme_bw() +
#   theme(axis.title = element_text()) +
#   ylab("Bike Rentals") +
#   xlab("Visibility") +
#   ggtitle("Bike Rental Volume By Visibility")+
#   theme(plot.title=element_text(hjust=0.5))+
#   guides(color='none')
# ##################################################################################
# ggRentalsBysnow <- nyc %>%
#   ggplot(aes(y = n,
#                  x = snow,
#                  color = weekday)) +
#   geom_point(show.legend = TRUE) +
#   theme(legend.position=c(0.95,0.95))+
#   theme(legend.title = element_text(face = "bold"))+
#   theme(legend.text = element_text(size = 8, colour = "black"))+
#   theme(legend.key = element_rect(fill = "white", colour = "grey64"))+
#   facet_grid(~weekday) +
#   scale_color_brewer(palette = "Dark2") +
#   theme_bw() +
#   theme(axis.title = element_text()) +
#   ylab("Bike Rentals") +
#   xlab("Snow") +
#   ggtitle("Bike Rental Volume By Snow")+
#   theme(plot.title=element_text(hjust=0.5))+
#   guides(color='none')
# grid.newpage()
# ly <- grid.layout(2, 3)
# print(ggRentalsBycloudcover, vp = viewport(layout.pos.row = 1,
#                        layout.pos.col = 1))
# print(ggRentalsByhumidity, vp = viewport(layout.pos.row = 2,
#                        layout.pos.col = 1))
# print(ggRentalsByprecip, vp = viewport(layout.pos.row = 3,
#                        layout.pos.col = 1))
# })

```



## Impact of Weather Conditions on Bike rides

```{r}




vp1 <- viewport(x = 0, y = 0.5, width = 1, height = 0.5,
                just = c("left", "bottom"))
vp2 <- viewport(x = 0, y = 0, width = 1, height = 0.5,
                just = c("left", "bottom"))


draw_plot <- function(var_to_filter_by) {
  a <- required_vars %>%
    filter(var == !!var_to_filter_by) %>%
  ggplot(aes(x = value, y = speed, color = weekday)) +
  geom_point(size = 3) +
 # facet_wrap(~ var, scales = "free") +
  theme_bw()+
  theme(axis.title = element_text(size = 10)) +
  xlab(as.character(var_to_filter_by))+
  ylab("Average Speed of bikers (mph)") +
  ggtitle(paste0("Impact of ", as.character(var_to_filter_by) ," on Average Speed"))+
  theme(plot.title = element_text(face = "bold",hjust = 0.5,color="grey20"))
  
  b <- cnt_bikers_weather %>%
    filter(var == !!var_to_filter_by) %>%
    ggplot(aes(x = value, y = n, color = weekday)) +
  geom_point(size = 3) +
  #facet_wrap(~ var, scales = "free") +
  theme_bw()+
  theme(axis.title = element_text(size = 10)) +
  xlab(as.character(var_to_filter_by)) +
  ylab("Number of bikers") +
  ggtitle(paste0("Impact of ", as.character(var_to_filter_by)," on biker density"))+
  theme(plot.title = element_text(face = "bold",hjust = 0.5,color="grey20"))
  
  grid.newpage()
print(a, vp = vp1)
print(b, vp = vp2)
    
}

ui <- fluidPage(
  inputPanel(
    selectInput(
      "var",
      label = "Select Weather Condition",
      choices = required_lst
    )
  ),
  plotOutput('reqplot')
)

server <- function(input, output) {
  output$reqplot <- renderPlot(draw_plot(input$var), height = 325, width = 750)
}

shinyApp(ui = ui, server = server)



```

## THANK YOU!!




